---
title: "SQL babynames presentation"
author: "Brennan Donnell"
date: "10/9/2018"
output: ioslides_presentation
---
## Introduction: 
Data exploration in SQL and R on the *Babynames* data set. 

(Also, an exploration in *ioslides_presentation* in RMarkdown)
This is still a work in progress, and due to frame limitations, I can't show everything. 

```{r libraries&login, warning=FALSE, message=FALSE, error=FALSE, echo=FALSE}
library(data.table)
library(stringr)
library(ggplot2)
library(plotly)
library(dplyr)
library(dbplyr)
library(DBI)
library(RMariaDB)
library(devtools)
library(odbc)
library(readr)
library(RODBC)
uname <- 'bdonnell'
pword <- '1864'
df <- fread('/Users/brennandonnell/projects/GitHub/SQL-babynames-project/fullnames.csv', sep=',', 
            drop=1, colClasses = list(character=1:2, numeric=3:4))
df$length <- str_length(unlist(df[[1]]))
```

## Prelude to SQL: making a big file
Before loading all of the data into MariaDB, I had to merge it all together and make a column for year. 
I accomplished this in the following cell. Source: [Stack Overflow](https://stackoverflow.com/questions/41006985/importing-multiple-csv-files-into-r-and-adding-a-new-column-with-file-name)
```{r modcols_import, eval=FALSE, warning=FALSE, message=FALSE,error=FALSE}
path <- 
  '/Users/brennandonnell/projects/GitHub/SQL-babynames-project/'
file_names <- 
  dir(path, 
      pattern = "yob" ) 
df <- do.call(
  rbind, 
  lapply(file_names, function(x) cbind(fread(
           x, sep=',', 
           col.names = c('babyname','m/f','frequency'),
           colClasses = list(character=1:2, numeric=3)), 
           year=strsplit(strsplit(x,'yob')[[1]][[2]], '.txt')[[1]])))
df$length <- str_length(unlist(df[[1]]))
write.csv(df, paste0(path,'/fullnames2.csv'), sep=',')
```


## Loading mariadb in
Loading Mariadb into mySQL and R. 
This uses several libraries: 

>- DBI
>- dplyr
>- dbplyr
>- RMariaDB
>-  odbc

```{r load_mariadb}
con <- dbConnect(
  drv = RMariaDB::MariaDB(), 
  dbname = 'Babynames',
  username = uname,
  password = pword
)
```

## Creating the schema
this is just the framework for how to create a table in SQL. 
How I actually did it was by importing the full CSV into Sequel Pro through *file -> import*
```{sql, connection=con, max.print=5, tap.cap = "Creating a table"}
CREATE TABLE demonoo0 (
  baby_name CHAR(15),
  gender CHAR(1),
  frequency INT(11),
  year INT
)
```

## Query 1

For Query 1, we searched for the most popular male name and female name for that person’s year of birth. The following cell is a small sample of it. 
```{r mariadb_query1}
rs <- dbSendQuery(con, 
"SELECT babyname,gender, MAX(frequency), year 
FROM newnames 
GROUP BY gender, year 
ORDER BY year;")
head(dbFetch(rs))

dbClearResult(rs)
#dbDisconnect(con)
```

## Query 1: the SQL
It is possible in R Markdown to create chunks of SQL code. This is the same code as the previous slide, just done directly in SQL. 
```{sql, connection=con, max.print = 3, tab.cap="Most popular name by year"}
SELECT babyname,gender, MAX(frequency), year 
FROM newnames 
GROUP BY gender, year 
ORDER BY year;
```


## Query 2

For Query 2, we searched for the popularity of a given person’s name for each year from 1879 to 2017. 

Naturally, I picked my own *perfect* name, to see how many people have the same name as mine. 

```{r mariadb_query2}
rs <- dbSendQuery(con, 
"SELECT babyname, gender, frequency, year 
FROM newnames WHERE babyname = 'brennan' AND gender = 'm' 
ORDER BY year;")
tail(dbFetch(rs), n=3)
dbClearResult(rs)
```

## Query 2: SQL
Again, R Notebooks and R Markdown have the ability to run SQL code. This is the same code as the cell above. 
```{sql, connection=con, max.print = 3, tab.cap="Popularity of Brennan by year"}
SELECT babyname, gender, frequency, year FROM newnames 
WHERE babyname = 'brennan' AND gender = 'm' 
ORDER BY year;
```

## Plotting: 
For the following trick, we shall examine some visualizations. But first, we have to group the data. 
```{r plot_1_group}
by_brennan <- df %>%
  group_by(year) %>% 
  filter(babyname == "Brennan") 
```

## Plotting:
```{r plot_1_plot, message=FALSE, echo=FALSE}
p <- ggplot(data=by_brennan, aes(x=by_brennan$year, y=by_brennan$frequency, fill=by_brennan$`m/f`)) + geom_bar(stat='identity', 
                                                                                                                #fill="lightblue",
                                                                                                                position=position_dodge())
# p <- p+scale_color_manual(values=c("#E69F00", "#56B4E9"))
p <- p + labs(title='Popularity of my name by year', x='Year', y='Frequency')
p <- p + theme_minimal()
ggplotly(p)
```

## More plotting
look at name lengths
```{r plot_2_lengths}
by_length <- df %>% 
  filter(length >= 10) %>% 
  group_by(year) %>% 
  summarise(n = n())
```

## more plotting
```{r plot_2_plot, message=FALSE, echo=FALSE}
p <- ggplot(data=by_length, aes(x=by_length$year, y=by_length$n)) + geom_bar(stat='identity', fill="lightblue", position=position_dodge())
#p <- p+scale_color_manual(values=c("#E69F00", "#56B4E9"))
p <- p + labs(title='Number of names with at least ten letters', x='Year', y='Number of names')
p <- p + theme_minimal()
ggplotly(p)
```


```{r conclose, echo=FALSE}
dbDisconnect(con)
```